---
title: "Fire Locator"
author: Song Kim
date: September 9, 2020
output: ioslides_presentation
font-import: https://fonts.googleapis.com/css2?family=Roboto+Slab
font-family: 'Roboto Slab'
css: style.css
---

## Overview

Locate wildfires in the United States and display their locations and rate of growth

## Data

Wildfire data from the National Interagency Fire Center ([raw](https://opendata.arcgis.com/datasets/68637d248eb24d0d853342cba02d4af7_0.csv))

App searches for fires starting Jan. 1, 2020 (raw dataset starts in 2009, but data collection was spotty)

Input date search range

```{r data, warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(ggpubr)
download.file("https://opendata.arcgis.com/datasets/68637d248eb24d0d853342cba02d4af7_0.csv", "fire.csv", "curl")
fires <- read.csv("fire.csv", as.is=TRUE)
fires <- as.data.frame(rename(fires, lat=Y, long=X))
```

## Functions

Map shows locations, times of discovery, and daily growth in acres for each fire within the specified date range

Graphs show distribution of fire size (by rate of growth) and trend in the number of fires discovered each day

(Code below evaluated on next slide)
```{r processing, echo=FALSE}
fires <- fires[, c(2,1,3:dim(fires)[2])]
fires$fddate <- as.Date(fires$FireDiscoveryDateTime, "%Y/%m/%d %H:%M:%S")
fires$DailyAcres[is.na(fires$DailyAcres)] <- 0 # set NA acre increases to 0
fires$POOCity[is.na(fires$POOCity)] <- "" # set NA city names to blank
fires$POOState[is.na(fires$POOState)] <- "" # set NA state names to blank
```
```{r graphs, warning=FALSE, message=FALSE}
freqplot <- qplot(fddate, data=fires, xlab="Date of Discovery", ylab="Count")
acreplot <- ggplot(data=fires, aes(fddate, DailyAcres)) + 
  geom_point() + scale_x_date(date_breaks="6 months", date_labels="%b %y") + 
  labs(x="Date of Discovery", y="Acres Daily")
figure <- ggarrange(freqplot, acreplot + font("x.text", size = 10),
                    ncol = 1, nrow = 2)
```

## Results

There are not too many fires until about the beginning of September. Around September 6/Sept. 7, the daily fire discovery rate increased significantly.
```{r show, echo=FALSE}
figure
```